---
title: "index"
author: "Ana Sofia Guerra"
date: "November 30, 2017"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Scatterplot Party! trying again

```{r mpg, include=FALSE}
library(tidyverse)
#mpg
```

```{r scatterplot}
ggplot(data=mpg, aes(x=displ, y=hwy)) + geom_point()
```

### I need `color`!

```{r color scatter}
ggplot(data=mpg, aes(x=displ, y=hwy, color=class)) + geom_point()
```

...but I need a black and white figure, so change `color` to `shape`

```{r bw scatter}
ggplot(data=mpg, aes(x=displ, y=hwy, shape=class)) + geom_point()
```

Keep shapes but need to change the size of shapes and add in fuel type as a color and now looking at cty instead of displ

`alpha` in `geom_point` adds transparency 

```{r}
ggplot(data=mpg, aes(x=cty, y=hwy, size=class, color=fl)) + 
  geom_point(alpha=0.2, color= "blue")
```



```{r}
    ggplot(data = mpg, aes(x = displ, y = hwy, olor=displ <5)) + 
      geom_point()
```

###Geoms

```{r}
ggplot(mpg, aes(x=drv, y=hwy))+
  geom_boxplot()
```

```{r}
ggplot(mpg, aes(x=drv, y=hwy))+
  geom_jitter()
```
```{r}
ggplot(mpg, aes(x=drv, y=hwy))+
  geom_violin()
```


combining two types of geoms into one
```{r}
ggplot(data=mpg, aes(x=displ, y=hwy)) + 
  geom_point()+
  geom_smooth()
```
Now lets customize this plot

color funtimes

```{r palettes}
install.packages("colorspace")
library(colorspace)
```

```{r customize}
ggplot(data=mpg, aes(x=displ, y=hwy, color=cyl)) + 
  geom_point(alpha=0.9, size=4)+
  geom_smooth(fill="pink", color="red",size=1)+
  labs(title="Relationship between engine size and mpg", y="Highway mpg", x="Enginge displacement (liters)")+
  theme_classic()+
  theme(text=element_text(size=15), panel.grid.major=element_blank())+
  scale_color_continuous(name="# Cylinders")
```
##Data Wrangling with `dplyr`

get gapminder dataset from OHI website

```{r}
gapminder <- read_csv("https://raw.githubusercontent.com/OHI-Science/data-science-training/master/data/gapminder.csv")
```

`read_csv` is from the `readr` package (part of tidyverse). It is not `read.csv`, which is part of base R. 

###useful functions to get to know data

```{r}
head(gapminder) #first six rows of data
tail(gapminder) # last six rows of data

head(gapminder, 30) #show X n rows of data instead of the default six

str(gapminder) #structure
names(gapminder) #colnames
dim(gapminder) #dimensions
ncol(gapminder) #number of columns
nrow(gapminder) #number of rows

c(nrow(gapminder), ncol(gapminder)) #c() helps you combine two different functions
summary(gapminder)
```

For everything above we were operating on the entire gapminder dataset. 

```{r}
head(gapminder$lifeExp)
```

###Actually working with diplyr 

Using `filter`
```{r}
filter(gapminder, lifeExp <29) #from gapminder data, filters out those where life expectancy is less than 29

filter(gapminder, country == "Mexico")

filter(gapminder, country == "Sweden") %>% summarise((mean(lifeExp)))

```

####Pipe operator `%>%` or shortcut Cmd+Shift+M, pipe reads as THEN

```{r}
gapminder %>% head() #instead of head(gapminder)ss

gapminder %>% head(3) #still lets you decide how many rows in head. 

gapminder %>% filter(country =="Sweden") %>% summarise(mean(lifeExp))
```

###Select() by columns
```{r}
select(gapminder, year, lifeExp)

gapminder %>% select(year,lifeExp) %>% tail()
```

###combine all the things


```{r}
gapminder %>% filter(country=="Cambodia") %>% select(country, year, pop, gdpPercap)

gapminder %>% filter(country=="Cambodia") %>%  select(-continent,-lifeExp) #exclude/deselect rather than select for specific columns

gapminder %>% 
  filter(country=="Mexico", 
         year==2002)
```

###mutate to add columns

```{r}
gapminder %>% mutate(gdp = pop*gdpPercap)
```
#####let's add an index

```{r}
test <- gapminder %>% 
  mutate(gpd = pop*gdpPercap) %>% 
  mutate(index= 1:nrow(gapminder)) 
```

ilter for multiple values
`%in%` lets you filter multiple things within a `c()`

```{r}
gapminder %>% 
  filter(country %in% c("Egypt", "Vietnam")) 
```

Exercise: max gdpPerCapita of Egypt and Vietnam in new column 

```{r}
gapminder %>%
  filter(country %in% c("Egypt", "Vietnam")) %>%
  mutate(maxgpdpc = max(gdpPercap))
```

####group_by() so that we can get two max values

```{r}
gapminder %>% 
  filter(country %in% c("Egypt", "Vietnam")) %>% 
  group_by(country) %>% 
  summarise(max_gdpPercap = max(gdpPercap))
```
If you don't specify which country, it gives max for all countries
```{r}
gapminder %>% 
  group_by(country) %>% 
  summarise(max_gdpPercap = max(gdpPercap))
```

##let's keep the year associated with that max_gdpPerCap

```{r}
gapminder %>% 
  group_by(country) %>% 
  mutate(max_gdpPercap = max(gdpPercap)) %>% 
  filter(max_gdpPercap == gdpPercap) %>% 
  arrange(max_gdpPercap) #arrange lets you order the data in descending order. 
```

